**Selective DNA and Protein Isolation from Marine Macrophyte Surfaces**

### 2020-11-19 submission/manuscript.Rmd
Format customization.

\newpage
[]: # (\linenumbers)
\sisetup{mode=text}
\setlength\parindent{24pt}

@@

## Discussion
To test whether the developed DNA and protein isolation protocols efficiently detach microbes from the macrophyte surface we selected *C. nodosa* and *C. cylindracea* as representatives of seagrass and macroalgal species. These species differ morphologically. While *C. nodosa* leaves are flat, *C. cylindracea* thalli are characterized by an uneven surface [@Kuo2001; @Verlaque2003]. The developed protocol led to an almost complete removal of epiphytic cells from the surfaces of both species comparable to the result of @Burke2009, indicating that structural differences do not impact the removal efficiency. In addition, isolation protocols were tested in two contrasting seasons, as it is known that macrophytes are harbouring more algal epiphytes during autumn and winter [@Reyes2001]. No differences in the removal efficiency was observed between seasons suggesting that these protocols can be used on macrophyte samples retrieved throughout the year. Also, no removal differences were observed on samples derived from the same host but from different locations.

@@

### Illumina 16S rRNA sequencing
An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure. In the first PCR the 515F (5\prime-GTGYCAGCMGCCGCGGTAA-3\prime) and 806R (5\prime-GGACTACNVGGGTWTCTAAT-3\prime) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016]. These primers contained on their 5\prime end a tagged sequence. Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield. The column was eluted in 30 \si{\ul} of deionized water. Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany. Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR. In addition, these primers contained adapter and sample-specific index sequences. The second PCR was carried out for 8 cycles. Beside samples, a positive and negative control were sequenced. A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used. Partial 16S rRNA sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811.

>>>>

\newpage
\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}
\doublespacing

@@

\newpage
## Discussion
To test whether the developed DNA and protein isolation protocols efficiently detach microbes from the macrophyte surface we selected *C. nodosa* and *C. cylindracea* as representatives of seagrass and macroalgal species. These species differ morphologically. While *C. nodosa* leaves are flat, *C. cylindracea* thalli are characterized by an uneven surface [@Kuo2001; @Verlaque2003]. The developed protocol led to an almost complete removal of epiphytic cells from the surfaces of both species comparable to the result of @Burke2009, indicating that structural differences do not impact the removal efficiency. In addition, isolation protocols were tested in two contrasting seasons, as it is known that macrophytes are harbouring more algal epiphytes during autumn and winter [@Reyes2001]. No differences in the removal efficiency was observed between seasons suggesting that these protocols can be used on macrophyte samples retrieved throughout the year. Also, no removal differences were observed on samples derived from the same host but from different locations.

@@

### Illumina 16S rRNA sequencing
An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure. In the first PCR the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016]. These primers contained on their $5'$ end a tagged sequence. Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield. The column was eluted in 30 \si{\ul} of deionized water. Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany. Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR. In addition, these primers contained adapter and sample-specific index sequences. The second PCR was carried out for 8 cycles. Beside samples, a positive and negative control were sequenced. A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used. Partial 16S rRNA sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811.
